Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2019-12-09 xrb/paper.tex
Grammar

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \rho \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the
pressure and $E$ is the specific total energy, which is related to the
specific internal energy as $e = E - |\Ub|^2/2$.  The forcing in the
momentum equation includes gravity, described by gravitation
acceleration $\gb$ and rotational forces described by angular velocity
$\Omegab$, with $\rb$ the position vector from the origin.
Species are described by mass fractions, $X_k$ (such that $\sum_k X_k
= 1$), and creation rates, $\omegadot$, and are related to the total
specific energy generation rate, $\epsdot$.  The total mass
conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  An equation of state of the form $p = p(\rho, e, X_k)$ completes the
thermodynamic description of the system.  Thermal diffusion is 
described by a thermal conductivity $\kth$ and temperature $T$.

@@

To understand the time and length scales involved in flame
propagation, we do a 1D simulation of a laminar flame using our
microphysics.  Figure~\ref{fig:flame} shows the flame thermodynamic
profile and properties for the 10/10 flame using
% \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
our conductivities and the {\tt aprox13} reaction network.  This flame
had a density of $2\times 10^6~\gcc$ and
temperature of $5\times 10^7$~K.  We observe that this flame speed is
about $10^5~\cms$, the flame width is about 40~cm, and it
takes about 3~ms to settle into a sustained flame.  Note that this speed
is quite small compared to the speeds of $\sim 10^6~\cms$ estimated in
\citet{spitkovsky2002}.  Table~\ref{table:flame_speeds_1d} gives the properties
of the 10/10 and 5/5 laminar flames.  We expect a multidimensional flame to
accelerate due to hydrodynamics interactions (wrinkling, turbulence
interactions, directed flows feeding fuel into the flame, ...).

@@

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean-molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}.
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonably well-defined.  We see that the bottom of the burning
front is lifted off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect, as we'll see in the 5/5 case below.

>>>>

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \rho \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the
pressure and $E$ is the specific total energy, which is related to the
specific internal energy as $e = E - |\Ub|^2/2$.  The forcing in the
momentum equation includes gravity, described by gravitational
acceleration $\gb$, and rotational forces, described by angular velocity
$\Omegab$, with $\rb$ the position vector from the origin.
Species are described by mass fractions, $X_k$ (such that $\sum_k X_k
= 1$), and creation rates, $\omegadot$, and are related to the total
specific energy generation rate, $\epsdot$.  The total mass
conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  An equation of state of the form $p = p(\rho, e, X_k)$ completes the
thermodynamic description of the system.  Thermal diffusion is 
described by a thermal conductivity $\kth$ and temperature $T$.

@@

To understand the time and length scales involved in flame
propagation, we do a 1D simulation of a laminar flame using our
microphysics.  Figure~\ref{fig:flame} shows the flame thermodynamic
profile and properties for the 10/10 flame using
% \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
our conductivities and the {\tt aprox13} reaction network.  This flame
had a density of $2\times 10^6~\gcc$ and
temperature of $5\times 10^7$~K.  We observe that this flame speed is
about $10^5~\cms$, the flame width is about 40~cm, and it
takes about 3~ms to settle into a sustained flame.  Note that this speed
is quite small compared to the speeds of $\sim 10^6~\cms$ estimated in
\citet{spitkovsky2002}.  Table~\ref{table:flame_speeds_1d} gives the properties
of the 10/10 and 5/5 laminar flames.  We expect a multidimensional flame to
accelerate due to hydrodynamics interactions (wrinkling, turbulence
interactions, directed flows feeding fuel into the flame, etc.).

@@

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}.
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonably well-defined.  We see that the bottom of the burning
front is lifted off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect, as we'll see in the 5/5 case below.
